Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey

### 2014-01-26 intro.tex
2012.10.08 (Vassilis) Text added explaining the three indicators (low, medium, high) for each of the criteria against which the algorithms are evaluated (efficiency, overhead, scalability). Filled in missing references and question marks in the tables and text. Intro and background reworked. First page shows the full abstract and bar charts calibrated (orientation, axis labels etc). Text now conforms to ACM Computing Surveys style/format.

Peer-to-peer (P2P) networks are self-organizing distributed systems
where participating nodes, called \emph{peers}, act as both resource
providers and resource consumers, in contrast to the traditional \emph{client-server}
model where nodes take on the role of pure servers or pure clients.  
Peer-to-peer networks have enjoyed immense attention from Internet users,
companies, and researchers and have been widely
deployed for over a decade, primarily because of the great
number of features they offer distributed applications that are built
on top of them.  These features include:  high availability and robustness,
load-balancing, quality of service, scalability, decentralized
administration, and anonymity.

A myriad of applications have been built and deployed using the peer-to-peer
paradigm including two widely-known ``killer'' applications:  file-sharing and telephony.
P2P file-sharing was originally introduced by Napster (see Section 2)
which by 2001 was widely acknowledged as the ``fastest growing Internet application
ever" with 26 million users sharing 80 million songs at the height of its popularity.  
Other file-sharing applications
quickly followed suit including Gnutella and its descendent, Limewire, connecting
3 million users concurrently by the end of 2006 and BitTorrent, connecting
over 150 million users.
P2P telephony
has seen explosive growth with the advent of Skype.  Introduced in 2003, Skype
has experienced explosive popularity with 
more that 650 million users in 2011 \cite{skypetotalusers} and
35 million skypers concurrently online on
the 5th of March 2012 \cite{skypesymusers}.

In addition to file-sharing and telephony, a large number of diverse applications have
been built on top of peer-to-peer systems including:
distributed search engines \cite{yaci}, distributed data storage systems
\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003}, Web caches, archives,
and publishing systems \cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
messaging applications \cite{threedegrees}, event notification infrastructures
\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, naming services
\cite{cmm_chorddns_2002}, censor-resistant stores \cite{cswh_freenet_2001} and
lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.

Peer-to-peer systems are implemented using overlay networks.   
An overlay network is a virtual network of nodes and logical links 
that is built on top of an existing network with the aim of providing 
a service (such as content distribution) that is not available 
in the existing network.  Nodes in the overlay are connected via virtual 
links, each of which corresponds to a path of possibly many physical 
links in the underlying network.  
Figure~\ref{figure:overlay}
illustrates a simple four-node overlay constructed over a wide-area network.
In the case of peer-to-peer networks, typically the participating nodes
are user PCs sitting at the edge of the Internet.

@@

One of the fundamental issues that defines the efficiency of an overlay 
network is how 
well the overlay network topology maps to the underlying network topology on which it rests.
Consider two nodes that share a link in an overlay network which maps to multiple 
hops in the underlying IP topology.  If the traffic patterns of the application are such
that traffic flows heavily over the overlay link between the two nodes, then it is
beneficial to construct the overlay topology such that the number of underlying IP links
between these two nodes is minimized.

If the overlay network is constructed such that the topology does not match the 
underlying IP topology well, the resulting \emph{topology mismatch} 
can result in
two significant problems.  First, application performance (e.g., search query latency
or jitter) can be adversely affected since application traffic must flow over a larger number
of IP hops than if the overlay were constructed more wisely.  Second, studies have shown
that P2P traffic contributes the largest portion of overall Internet traffic
\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
some reporting that more than 60\% of the traffic is P2P~\cite{cachelogic,ipoque2007,ipoque2009}  
and others reporting that P2P traffic is expected to reach almost 8 petabytes 
per month by the end of 2012~\cite{multinteligence}.  This is a significant burden for Internet Service Providers (ISPs) who 
must transfer this traffic to destination nodes sitting at the edge of the Internet.
If the overlay topology of a peer-to-peer system is not cleverly constructed, then 
the burden of ISPs carrying P2P traffic could be even greater.
For example, traffic might have to
flow back and forth several times between two neighboring ISPs as it travels from its
source node in the overlay to its destination node.
Thus, it is fundamentally important that P2P overlay networks are constructed
and maintained so that
their topology matches as closely as possible the underlying IP topology.

For over a decade, several tens of researchers have published profusely on the 
topology mismatch problem.  In this paper, we survey the literature
on this topic and provide a taxonomy of the proposed approaches and solutions.
We point out synergies
as well as similarities and differences in the published approaches.  
Ultimately, our goal is to help readers sift through the vast
amount of literature, to help them understand the advantages and 
disadvantages of each approach, and to provide them with enough 
perspective so that when the need arises, they are able to select 
amongst the different approaches for their 
particular application.

The rest of the paper is as follows:  In Section~\ref{section:background}, 
we provide some background information
on peer-to-peer overlay architectures including centralized, decentralized unstructured
and decentralized structured peer-to-peer systems.  We also present formally the
problem of alleviating the topology mismatch between a peer-to-peer overlay and the
underlying IP network.  In Sections~\ref{section:unstructured} and 
\ref{section:structured}, we describe a taxonomy of
the research efforts and solutions
proposed for decentralized unstructured peer-to-peer systems 
and decentralized
structured peer-to-peer systems, respectively.  Finally, 
in Section~\ref{section:conclusion}, we conclude.

>>>>

\emph{Peer-to-peer (P2P)} networks are self-organizing distributed systems where
participating nodes, called \emph{peers}, act as both resource providers and
resource consumers, in contrast to the traditional \emph{client-server} model
where nodes take on the role of pure servers or pure clients. Peer-to-peer
networks have enjoyed immense attention from Internet users, companies, and
researchers and have been widely deployed for over a decade, primarily because
of the great number of features they offer distributed applications that are
built on top of them. These features include:  high availability and robustness,
load-balancing, quality of service, scalability, decentralized administration,
and anonymity.

Two ``killer'' applications of the peer-to-peer paradigm are, but not limited
to, file-sharing and telephony. P2P file-sharing was originally introduced by
Napster (see Section~\ref{section:background}), which by 2001 was widely
acknowledged as the ``fastest growing Internet application ever" with 26 million
users sharing 80 million songs at the height of its popularity. Other similar
applications quickly followed suit including Gnutella and its descendent,
Limewire, enjoying 3 million concurrently connected peers by the end of 2006
and BitTorrent, connecting over 150 million users. P2P telephony has seen
explosive growth with the advent of Skype. Introduced in 2003, Skype has
experienced explosive popularity with more that 650 million users in 2011
\cite{skypetotalusers} and 35 million skypers concurrently online on the 5th of
March 2012 \cite{skypesymusers}.

In addition to file-sharing and telephony, a large number of diverse
applications have been built on top of peer-to-peer systems including:
distributed search engines \cite{yaci}, distributed data storage systems
\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003},
Web caches, archives and publishing systems
\cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
messaging applications \cite{threedegrees}, event notification infrastructures
\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, naming services
\cite{cmm_chorddns_2002}, censor-resistant stores \cite{cswh_freenet_2001} and
lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.

Peer-to-peer systems are implemented using overlay networks. An \emph{overlay
network} is a virtual system of nodes and logical links, that is built on top of
an existing network with the aim of providing a service (such as content
distribution) that is not available in the existing one. Nodes in the overlay,
communicate via virtual connections, each of which may correspond to a path of
possibly many physical links in the underlying network.
Figure~\ref{figure:overlay} illustrates a simple four-node overlay constructed
over a wide-area network.

@@

One of the fundamental issues that defines the efficiency of an overlay network
is how well the overlay maps to the underlying network topology on which it
rests. Consider two nodes\footnote{In the case of peer-to-peer networks,
typically the participating nodes are user PCs sitting at the edge of the
Internet.} that share a link in an overlay network which maps to multiple hops
in the underlying (say IP) topology. If the traffic patterns of the application
are such that traffic flows heavily over the overlay link between the two nodes,
then it is beneficial to construct the overlay topology such that the number of
underlying IP links between these two nodes is minimized.

If the overlay network is constructed such that the topology does not match the 
underlying IP topology well, the resulting \emph{topology mismatch} (See
Section~\ref{section:background} for a more formal definition) can result in two
significant problems. First, application performance (e.g. search query latency
or jitter) can be adversely affected since application traffic must flow over a
larger number of IP hops than if the overlay were constructed more wisely.
Second, studies have shown that P2P applications contribute the largest portion
of overall Internet traffic
\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
some reporting that more than 60\% of it, is
P2P-related~\cite{cachelogic,ipoque2007,ipoque2009} and others claiming that it
is expected to reach almost 8 petabytes per month by the end of
2012~\cite{multinteligence}. This is a significant burden for Internet Service
Providers (ISPs) who must handle all this traffic to destination nodes sitting
at the edge of the Internet. If the overlay topology of a peer-to-peer system is
not efficintly constructed, then this burden to the Internet's backbone
infrastructure could be even greater. For example, traffic might have to flow
back and forth several times between two neighboring ISPs as it travels from its
source node in the overlay to its destination node. Thus, it is fundamentally
important that P2P overlay networks are constructed and maintained so that
their topology matches as closely as possible the underlying IP topology.

For over a decade, several tens of researchers have published profusely on the 
topology mismatch problem.  In this paper, we survey this research area and
provide a taxonomy of the proposed solutions. We point out synergies as well as
similarities and differences in the published approaches. Ultimately, our goal
is to help readers sift through the vast amount of literature, to help them
understand the advantages and disadvantages of each of each work, and to provide
them with enough perspective so that when the need arises, they are able to
select amongst the different approaches for their particular application.

The rest of the paper is as follows: In Section~\ref{section:background}, we
provide some background information on peer-to-peer overlay architectures
including centralized, decentralized unstructured and decentralized structured
peer-to-peer systems. We also formally define the problem of topology mismatch
between an overlay and its underlying physical network and elaborate on the
rationale behind our coarse grained evaluation of the techinques presented
throughout this survey. In Sections~\ref{section:unstructured} and
\ref{section:structured}, we present the various research efforts for
decentralized unstructured and decentralized structured peer-to-peer systems,
respectively. Finally, with Section~\ref{section:conclusion}, we
conclude.

### 2014-01-26 mard-survey.tex
2012.06.26 (Mema) Big introduction section split to a crispier 3-page intro and a background section. The latter also shortened and updated.
2012.07.12 (Vassilis) Merged introduction section. Sectioning according to the method an algorithm is using dropped (almost all use a combination of two or more). First attempt for the tabular data on an external file.
2012.08.20 (Vassilis) Added LaTeX tables (known issue on wrapping text inside cells). Reshaped discussion material. Introducing rough, algorithm performance, estimation (high, medium, low) by tracking efficiency, control overhead and scalab

% packages
\usepackage[utf-8]{inputenc}
\usepackage{paralist}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{color,soul}
\usepackage{longtable}
\usepackage{pdflscape}

@@

\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured
Overlay Networks: A Survey}

\author{
VASSILIS S. MOUSTAKAS\\
\texttt{b.moystakas@di.uoa.gr}\\
\emph{Department of Informatics and Telecommunications}\\
\emph{National and Kapodistrian University of Athens}\\
H\"USEYIN AKCAN\\
\texttt{huseyin.akcan@ieu.edu.tr}\\
\emph{Department of Software Engineering}\\
\emph{Izmir University of Economics}\\
MEMA ROUSSOPOULOS\\
\texttt{mema@di.uoa.gr}\\
\emph{Department of Informatics and Telecommunications}\\
\emph{National and Kapodistrian University of Athens}\\
AND\\
ALEX DELIS\\
\texttt{ad@di.uoa.gr}\\
\emph{Department of Informatics and Telecommunications}\\
\emph{National and Kapodistrian University of Athens}
}

@@

\category{}{}{}

\terms{Topology mismatch, distributed, peer-to-peer, P2P, overlay}

\keywords{Topology mismatch problem, distributed algorithm, peer-to-peer
architecture, overlay network }

@@

\begin{bottomstuff}
SAY SOMETHING COOL HERE...
\end{bottomstuff}

@@

\input{intro}

>>>>

% packages
\usepackage[utf-8]{inputenc}
\usepackage{paralist}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{color,soul}
\usepackage{longtable}
\usepackage{wasysym}
\usepackage{pdflscape}
%\usepackage{tikz}
\usepackage{pgfplots}

@@

\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured
Overlay Networks: A Survey}
% TODO: We added here "P2P". I'm not sure, because for example Narada is not P2P
% TODO: I deleted "P2P"

\author{
VASSILIS S. MOUSTAKAS$^1$, H\"USEYIN AKCAN$^2$, MEMA ROUSSOPOULOS$^1$ and ALEX
DELIS$^1$\\
$^1$\emph{National and Kapodistrian University of Athens}, \emph{Athens},
\emph{Greece}\\
$^2$\emph{Izmir University of Economics}, \emph{Izmir}, \emph{Turkey}
}

@@

\category{A.1}{General Literature}{Introductory and Survey}
\category{C.2.4}{Distributed Systems}{Distributed Applications}

\terms{Topology mismatch}

\keywords{Distributed systems, peer-to-peer architectures, overlay network,
topology awareness}

@@

\begin{bottomstuff}
Authors' addresses: Department of Informatics and Telecommunications, National
and Kapodistrian University of Athens, Athens, Greece. Department of Software
Engineering, Izmir University of Economics, Izmir, Turkey.\\
emails: \{b.moystakas,mema,ad\}@di.uoa.gr, huseyin.akcan@ieu.edu.tr
\end{bottomstuff}

@@

\input{intro}
\input{background}
